---
title: Evaluation of LC-MS contamination risk (HowDirty)
output:
  html_document:
    toc: true
    toc_float: false
    theme: united
    number_sections: TRUE
params:
  DataSet: 
    label: "Data set or experiment (keep it short since it's used as prefix for output file names)"
    value: 2021-231_FASP_filters
    input: text
  UserNames: 
    label: "User(s)"
    value: 
    input: text
  Notes:
    label:
    value:
    input: text
  PeakAreasContaminantsFile:
    label: "PeakAreaContaminants report from Skyline"
    value: data/PeakAreas_Contaminants_2021-231.csv # PeaKAreas_Contaminants report (directory/name) resulting from Skyline
  AnnotationFile:
    label: "Sample annotation file"
    value: data/samples_annotation_2021-231.csv # Samples Annotation directory/name
  RefThresholdsFile:
    label: "Reference threshold file (total)"
    value: data/2021-231_FASP_filters_report_contaminants_20230712_1821.xlsx # A file directory/name in or FALSE if no threshold available
  OutputDirectory:
    label: "Directory where the report and the results are exported"
    value: results
  RemoveMissingContaminants:
    label: "Remove contaminants that were not identified in any sample"
    value: TRUE
  nTopContaminantGroups:
    label: "Top n contaminant groups to show in plots"
    value: 10
    input: numeric
  MultiplyDilutionFactor:
    label:
    value: FALSE #Use carefully, normalization by TIC already takes into account slight differences in injection amounts
  PlotsInteractive: 
    label: "Export interactive plots"
    value: TRUE #TRUE = interactive in html, FALSE = static
---

<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

# Report information

## Data set information

**Data set:** `r params$DataSet`

**User(s):** `r params$UserNames`

**Date:**  `r Sys.Date()` 

**Notes:** `r params$Notes`

**Objective:** Evaluate the risk of possible contamination with polymers and small molecules in the samples from the data set.

## HowDirty evaluation of LC-MS results

This report was generated using *HowDirty*. This is meant to help you evaluating *How Dirty* are your samples (or system) in LC-MS analyses. The possible contaminants reported hereby could impact the quality of the results (e.g. contaminant IDs and reproducibility) and, in extreme cases, even damage the column or instruments. The raw data are previously analyzed using Skyline to extract LC-MS peaks possibly corresponding to contaminants. Then, a Risk level is assigned to each contaminant by comparing against the reference threshold (see Procedures for details). Finally, the Risk is similarly assessed at the Sample and whole Sample Set levels. These are indicative of the level of possible contamination.

**Code author:** David Gomez-Zepeda (HI-TRON)

**Version:** 0.04

```{r setup, include=FALSE, warning=FALSE}
#General parameters for knitr
require("knitr")
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
print(getwd())
```

```{r verify_files, include=FALSE, echo=FALSE}
# verify that parameters were filled
if(is.null(params$PeakAreasContaminantsFile))  stop("Please enter the PeakAreasContaminantsFile")
if(is.null(params$RefThresholdsFile)) stop("Please enter the RefThresholdsFile or assign it as FALSE")
# verify that files exist
filenames <- unlist(params[grepl("File" , names(params))])
filenames <- filenames[filenames != "FALSE"] # ignore if FALSE (e.g. for thresholds)
filecheck <- file.exists(filenames) 
if(any(!filecheck)){
   files_missing <-     data.frame(File=names(filenames[!filecheck] ), Entry=filenames[!filecheck] , row.names=NULL)
   files_missing$message <- paste0(files_missing$File, ": ", files_missing$Entry, "\n")
   files_missing
  stop("The input files below were not found. Please verify that they are in the right location: \n ", files_missing$message)
}
```


```{r initialize, message=FALSE, include=FALSE}
#Install required packages
##Test whether all required packages, install packages needed and install if needed
required_packages <- c('ggplot2', 'ggpubr', 'scales', 'plotly',
              'openxlsx', 'mgsub', 'DT', 'shiny', 'tidyverse', 'HowDirty')
# required_packages <- c(packages)
installed_packages <- required_packages %in% installed.packages()[,"Package"]
missing_packages <- required_packages[!installed_packages]
if ( length(missing_packages) > 0 ) install.packages(packages)
##Load libraries
invisible(lapply(required_packages, require, character.only = TRUE))
# names for output files
today_date <- gsub(patt ="-", rep="",paste0(gsub(patt ="-", rep="", Sys.Date()), "_", format(Sys.time(), "%H%M")))
filedir_report <- file.path(params$OutputDirectory,
                            paste0(params$DataSet, "_report_contaminants_", today_date, ".xlsx"))
filedir_sessioninfo <- gsub(filedir_report, patt = ".xlsx", rep = "_RSessionInfo.txt", filedir_report)
if(!dir.exists(params$OutputDirectory)) dir.create(params$OutputDirectory)
# report-specific functions
print_plot_or_plotly <- function(input, height_plotly = 450, width_plotly = 1000, interactive = params$PlotsInteractive){
  # wrapper function to use ggplotly when params$PlotsInteractive == TRUE
  if(interactive == TRUE){
    ggplotly(input, height = height_plotly, width = width_plotly)
    # layout_ggplotly_label_margin( x = -0.0, y = -0.05) #fixes label margin problem
  }else{
    if(interactive == FALSE){
      print(input)
    }
  }
}
```

# Procedure

## Data analysis

* Raw data was imported into Skyline to extract the features corresponding to possible contaminants (ToDo: add references)
* For this purpose, the Molecular Contaminant List template was used (Gomez-Zepeda *et al.*, 2023; modified from Rardin, 2018)
* The feature area of extracted ion chromatograms was processed using *HowDirty* to generate this report (Gomez-Zepeda *et al.*, 2023)

## Warning

The algorithm used for peak picking in Skyline is simple and based only on m/z and charge (z). Thus, some true contaminant peaks (e.g. peptides) could be incorrectly assigned to contaminants (i.e. false positives). Therefore, it is recommended to also look into the Skyline file to evaluate other factors, such as patterns of contaminant groups elution across the retention time.

## Calculations within *HowDirty*

* TICA = Total Ion Current Area
* Abundance (Normalized abundance) = TotalAreaMS1 / TICA
* TotalAbundance_ContaminantGroup = Sum (Abundance_ContaminantGroup) across all the contaminants in one ContaminantGroup for one sample
* Abundance_total = Sum (Abundance) across all the contaminants for one sample
* Contaminant-specific Risk level assessment was performed by comparing the Abundance of the possible contaminants in each one of the test samples (current data set) against thresholds previously extrapolated from a reference data set (~ 1000s runs). These thresholds are reported in `r filedir_report`--> ref_conta_tshd
* Sample level summary contaminant group assessment was performed by comparing the TotalAbundance_ContaminantGroup against the summed thresholds from each contaminant
* Sample level summary contaminant risk assessment was performed by comparing the Abundance_total against the sample-level quantile thresholds from the reference dataset
* These thresholds are reported in `r filedir_report`--> ref_conta_tshd_sample

# Input

```{r read_data}
# read files
## Skyline results of contaminant peak area and height
conta_raw <- read_conta_results(file_report_skyline = params$PeakAreasContaminantsFile, simplify_ContaminantGroup = TRUE)
## information about samples (annotation)
samples_annot <- read_samples_annotation(params$AnnotationFile) 
## check if all samples are included in both results and annotation file
check_samples_in_results(conta_raw, samples_annot)
# Load thresholds

if(params$RefThresholdsFile == FALSE){
  # set arbitrary thresholds
  ref_conta_tshd <- get_simple_thresholds_contaminant(conta_raw)
  ref_conta_tshd_sample <- get_simple_thresholds_sample()
  message("WARNING: Reference Threshold File was not provided. Thus, threshold were set at the same level for all the contaminants")
} else{ 
  # read thresholds from file
  ref_conta_tshd <-  read_conta_thresholds(params$RefThresholdsFile)
  ref_conta_tshd_sample <-  read_conta_sample_thresholds(params$RefThresholdsFile)
}
```

```{r annotate_results}
# Annotate contamination results with samples_annot, multiply by dilution_factor, assign risk level
conta <-
  conta_raw %>% 
  annotate_conta_samples(., samples_annot, # Annotate with samples_annot, multiply by dilution_factor
                         remove_missing = params$RemoveMissingContaminants, 
                         multiply_dilution_factor = params$MultiplyDilutionFactor) %>% 
  annotate_conta_thresholds(., ref_conta_tshd, Abundance)  %>%  # assign RiskLevel
  arrange_contagroup_levels(metric = "median") # arrange contaminant group levels based on overall median abundance
```

The data set contained the following:

```{r summary_input}
options(width = 1000)
count_factor_columns(conta) %>% 
  filter(!Variable %in% c("Risk", "RiskLevel")) %>% 
  mutate(Variable = paste0(Variable, "s")) %>% 
  kable("html")
```

```{r summaries}
#Summary of thresholds for each contaminant group
ref_conta_tshd_sum <-  ref_conta_tshd %>% 
  group_by(ContaminantGroup) %>% 
  summarise(across(starts_with("Tshd"), sum), .groups = "drop")

## Summary by Contaminant Group, needed to get top groups
conta_summ_contaminantgroup <- summarize_conta(conta, ContaminantGroup)
conta_summ_contaminant <- summarize_conta(conta, ContaminantGroup, Contaminant)

conta_summ_sample <-
  conta %>% 
  summarize_conta(Condition, Sample, ReplicateName) %>% 
  annotate_conta_thresholds_samples(ref_conta_tshd_sample) 

conta_summ_contaminantgroup_sample <-
  summarize_conta(conta, Condition, Sample, ReplicateName, ContaminantGroup)  %>% 
  annotate_contagroup_thresholds(ref_conta_tshd, Abundance_median)

conta_summ_condition <-
  summarize_conta(conta, Condition, ContaminantGroup, Contaminant) %>% 
  annotate_conta_thresholds(ref_conta_tshd, Abundance_median)

#Dataset level
conta_summ_sampleset <-
  summarize_conta_sampleset(conta)
risk_summ_sampleset <-
  conta_summ_sample %>%
  count(Risk, name = "Count") %>%
  mutate(Fraction = paste0(Count, "/", sum(Count)) ,
         Freq =  round(Count/sum(Count), 2),
         Perc = label_percent()(Freq),
         Tag = paste0(Risk, ": ", Perc, " (", Fraction, ")")) 
risk_summ_sampleset_verbose <- 
  risk_summ_sampleset %>% 
  summarise(Risk  =  paste0(Tag, collapse = "; "), .groups = "drop")
# Verbose summary of the data set
max_risklevel <- max(as.numeric(conta_summ_sample$RiskLevel))
summ_msg_options <- c("Something may be wrong, all Abundances were = 0"  ,
                      "You don't like getting soappy, your samples are super clean!",
                      "You are clean to go!",
                      "Be careful, you have some medium dirty samples",
                      "WARNING: You have some dirty samples!",
                      "WARNING: You have some very dirty samples!!!")
summ_msg <- summ_msg_options[max_risklevel]
summ_color_scale <- scales::viridis_pal(end=0.8, option =  "plasma", direction = 1)(7)
summ_color <- summ_color_scale[max_risklevel]
```

# Summary risk evaluation (total contamination) {.tabset}

* Overall status:  `r colorize_text(summ_msg, summ_color)`
* Evaluation based on the total contaminant abundance

## Global

* The piechart below shows the percentages of samples associated to each risk level
* The results were exported to `r filedir_report` --> risk_summ_sampleset

```{r plot_risk_summ_sampleset, fig.width= 5, fig.height=2.5}
plot_pie_risk_summ_sampleset <- plot_risk_summ_sampleset(risk_summ_sampleset)
plot_pie_risk_summ_sampleset
```

## Conditions boxplot

* Statistical difference was assessed by a Wilcoxon signed-rank test

```{r plot_conta_summ_condition_risk, warning=FALSE, fig.width=4, fig.height=2.5}
plot_conta_summ_condition_risk <-
  plot_condition_risk_total_boxplot(conta_summ_sample,
                                    compare_means = TRUE,
                                    method = "wilcox.test") 
plot_conta_summ_condition_risk # note: the results of the statistical test can't be shown as interactive plot with ggplotly
```


## Sample (ordered by name)

* The table can also be found in `r filedir_report` --> conta_summ_sample.
* RiskLevel = "1) Very Low (OK)", "2) Low (OK)", "3) Medium (Warning)", "4) High (Warning)", "5) Very High (DO NOT PROCEED)", "6) No threshold in reference"

```{r plot_conta_summ_sample_risk_bysample, warning=FALSE, fig.width= 10, fig.height=3}
dot_plot_sample_risk_total_bysample <-
  plot_sample_risk_total(conta_summ_sample, order_x = "Sample") +
  facet_wrap(~Condition, scales = "free_x") 
print_plot_or_plotly(dot_plot_sample_risk_total_bysample + rremove("xlab"))
```

## Sample (ordered by abundance)

* The table can also be found in `r filedir_report` --> conta_summ_sample.
* RiskLevel = "1) Very Low (OK)", "2) Low (OK)", "3) Medium (Warning)", "4) High (Warning)", "5) Very High (DO NOT PROCEED)", "6) No threshold in reference"

```{r plot_conta_summ_sample_risk_byabundance, warning=FALSE, fig.width= 10, fig.height=3}
dot_plot_sample_risk_total_byabundance <- 
  plot_sample_risk_total(conta_summ_sample, order_x = "Abundance") 
print_plot_or_plotly(dot_plot_sample_risk_total_byabundance + rremove("xlab"))
```

# Contaminant group {.tabset}

* y axis is ordered from highest to lowest abundance

## Condition dotplot

```{r plot_ContaminantGroupCondition, fig.width= 8, fig.height=5}
#dotplot
## ContaminantGroup 
dot_plot_conta_summ_condition_risk_contaminantgroup <-
  plot_condition_risk_contaminant(conta_summ_contaminantgroup_sample,
                              order_x = "Condition", order_y = "Abundance", show_zeros = FALSE)



print_plot_or_plotly(dot_plot_conta_summ_condition_risk_contaminantgroup)
```

## Sample dotplot

```{r plot_ContaminantGroupSample_abundance, fig.width= 10, fig.height=5}
#dotplot
## ContaminantGroup 
dot_plot_conta_summ_sample_risk_contaminantgroup <-
  # conta_summ_sample_risk %>% 
  conta_summ_contaminantgroup_sample %>% 
  plot_sample_risk_contaminant(order_x = "Sample", order_y = "Abundance", show_zeros = FALSE)


print_plot_or_plotly(dot_plot_conta_summ_sample_risk_contaminantgroup)
```

## Sample boxplot

```{r plot_abundance_contaminantgroup_sample, fig.width=6, fig.height=5}
plot_abundance_contaminantgroup_sample <- 
   plot_abundance(conta_summ_contaminantgroup_sample, ContaminantGroup, Abundance_median)

print_plot_or_plotly(plot_abundance_contaminantgroup_sample)
```

# Contaminants per sample{.tabset}

* A summary of the Abundance per ContaminantGroup was calculated and is reported in the annexed report (`r filedir_report` --> conta_summ_contaminantgroup).
* The Abundance of the contaminants are reported in the plots below for all the samples in the data set.
* The Pseudochromatograms represent the Abundance in function of the RentetionTime (Apex of the peak). Those can be useful to evaluate the presence of the usual patterns of polymers, i.e. from small to larger molecules; as well as the reproducibility of the RentetionTime across replicates.

## Abundance boxplot{.tabset}

```{r plot_ContaminantGroupArea, results='asis', fig.height=4, fig.width=6, warning=FALSE}
if(params$PlotsInteractive == TRUE){
  plot_abundance_contaminantgroup_list <-
  lapply(levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)],
         function(x){
           plot_abundance(filter(conta, ContaminantGroup == x), Contaminant, Abundance) %>%
             ggplotly(width = get_w(), height = get_h()*1.1) %>% 
             htmltools::tagList()#required to print ggplotly inside asis
         }
  )
  }else{if(params$PlotsInteractive == FALSE){
    plot_abundance_contaminantgroup_list <-
      lapply(levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)],
             function(x){
               plot_abundance(filter(conta, ContaminantGroup == x), Contaminant, Abundance) 
             }
      )
  }
  }

names(plot_abundance_contaminantgroup_list) <-
  levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)]

for (i in 1:length(plot_abundance_contaminantgroup_list)) {
  cat("### ", names(plot_abundance_contaminantgroup_list)[i],"\n")
  if(params$PlotsInteractive == TRUE){
    print(htmltools::tagList(plot_abundance_contaminantgroup_list[[i]])) #required to print ggplotly inside asis
  }else{if(params$PlotsInteractive == FALSE){
    print(plot_abundance_contaminantgroup_list[[i]])
  }
  }
  cat('\n\n')
}
```

## Pseudochromatograms (Abundance vs. RT){.tabset}

```{r plot_ContaminantGroupPseudoChromato, results='asis', fig.height=6, fig.width=10, warning=FALSE}


if(params$PlotsInteractive == TRUE){
  plot_seudochromatogram_contaminantgroup_list <-
  lapply(levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)],
         function(x){
           plot_pseudochromatogram(filter(conta, ContaminantGroup == x)) %>% 
             ggplotly(width = get_w(), height = get_h()*0.8) %>%
             # layout_ggplotly_label_margin( x = 0.05, y = -0.05) %>%
             htmltools::tagList() 
         }
  )
  }else{if(params$PlotsInteractive == FALSE){
  plot_seudochromatogram_contaminantgroup_list <-
  lapply(levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)],
         function(x){
           plot_pseudochromatogram(filter(conta, ContaminantGroup == x))
         }
  )
  }
  }


names(plot_seudochromatogram_contaminantgroup_list) <-
  levels(conta$ContaminantGroup)[c(1:params$nTopContaminantGroups)]

for (i in 1:length(plot_seudochromatogram_contaminantgroup_list)) {
  cat("### ", names(plot_seudochromatogram_contaminantgroup_list)[i],"\n")
  if(params$PlotsInteractive == TRUE){
    print(htmltools::tagList(plot_seudochromatogram_contaminantgroup_list[[i]])) #required to print ggplotly inside asis
  }else{if(params$PlotsInteractive == FALSE){
    print(plot_seudochromatogram_contaminantgroup_list[[i]])
  }
  }
  cat('\n\n')
}
```

## Table

* The following table shows the results of possible contaminants detected in each sample.
* An extended version of this table containing the Area and Total-Ion-Count-Area (TICA) can also be found in `r filedir_report` --> conta.
* RiskLevel = "1) Very Low (OK)", "2) Low (OK)", "3) Medium (Warning)", "4) High (Warning)", "5) Very High (DO NOT PROCEED)", "6) No theshold in reference"

```{r table_conta, echo=FALSE, warning=FALSE}
conta %>%
  select(ContaminantGroup, Contaminant, Condition, Sample, ReplicateName, RiskLevel, Abundance) %>%
  arrange(desc(Abundance)) %>%
  datatable(filter = 'top',
            options = list(pageLength = 10, autoWidth = TRUE)) %>%
  formatStyle('RiskLevel',
              backgroundColor = styleEqual(c(0, 1, 2, 3, 4, 5, 6),c(scales::viridis_pal(end=0.8, option =  "plasma", direction = 1)(7))))
```

# Appendix

* Result tables are exported in folder `r filedir_report`.

```{r echo=FALSE, warning=FALSE}
res_list<- list(
  risk_summ_sampleset = risk_summ_sampleset,
  conta_summ_sampleset = conta_summ_sampleset,
  conta_summ_sample = conta_summ_sample,
  conta_summ_condition = conta_summ_condition,
  conta_summ_contaminantgroup = conta_summ_contaminantgroup,
  conta_summ_contaminant = conta_summ_contaminant,
  conta=conta, 
  ref_conta_tshd = ref_conta_tshd,
  ref_conta_tshd_sample = ref_conta_tshd_sample
)
write.xlsx(res_list,
           file = filedir_report)
writeLines(capture.output(sessionInfo()), filedir_sessioninfo)
```

# References
- Gomez-Zepeda *et al.*. HowDirty. Under preparation. (2023).
- Rardin, M. J. Rapid Assessment of Contaminants and Interferences in Mass Spectrometry Data Using Skyline. J. Am. Soc. Mass Spectrom. 29, 1327–1330 (2018).
